---
title: "Covid91 vaccine study Final2024"
author: "Peter Dang"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

Is the vaccine effective in males?

The null hypothesis: the new preventive vaccine (Covid921) doesn't exhibit any effectiveness on males

The alternative hypothesis: Covid921 does exhibit effectiveness on males.

Cat~Cat

```{r}
Males = FinalData[FinalData$sex=='M',]
head(Males)
```

## Graphical results for males 

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")
```

According to the graphical results, The graph shows that the vaccine is effective in preventing COVID in males, because the amount of  males who got the vaccine and contracted COVID was lower than the men who got the placebo. 

## Numerical results for males 

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```

According to the data, we can see that 26.20% of men got the vaccine and contracted COVID, while 73.80% of them got the placebo and contracted COVID. 

## Inferential results for males 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Based on the inferential results, the p-value from the chi-squared test is 0.00000000000000022, which means this is less than 0.05, therefore, we reject the null hypothesis. The odds ratio from the fisher's exact test was 2.815. This means that the vaccine has a 2.8 chance of being effective in males. 

## Females

Null hypothesis: Covid921 vaccine doesn't exhibit its effectiveness on females. 

Alternative hypothesis: Covid921 vaccine does exhibit its effectiveness on females. 

```{r}
Females = FinalData[FinalData$sex=='F',]
head(Females)
```

## Graphical results for females 

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")
```

According to the graph, we can see that it shows that the vaccine is effective in women who got COVID compared to women who got the placebo. 

## Numerical results for females 

```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```

We can see that about 35.07% of females who got the vaccine contracted COVID. While 64.93% of females who got the placebo contracted COVID. This implies that the vaccine is effective in preventing COVID for women.

## Inferential results for females 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

According to the inferential results, we can see that the p-value from the chi-squared test is less than 0.05, therefore we reject the null hypothesis. For the odds ratio in the fisher's exact test is 1.856, meaning that that there is approximately a 1.86 chance that the vaccine will be effective in females. 


## LGBTQ

Null hypothesis: the vaccine doesn't exhibit its effectiveness on people who are LGBTQ.

Alternative hypothesis: the vaccine does exhibit its effectiveness on people who are LGBTQ.

```{r}
gay = FinalData[FinalData$LGBTQ=='gay',]
head(gay)
```


## Graphical results for LGBTQ

```{r}
barchartGC(~infected + treatment,data=gay)
barchartGC(~infected + treatment,data=gay, type="percent")
```

There is about 20% of people who are LGBTQ who got COVID and received the vaccine. This is lower than the amount of LGBTQ people who got COVID and received the placebo. The vaccine is effective in stopping COVID in people who are LGBTQ. 

## Numerical results for LGBTQ

```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

According to the data, we can see that 20% of LGBTQ people who got the vaccine contracted COVID, while 80% of LGBTQ people who got the placebo contracted COVID. Therefore we can say that the vaccine is effective for them. 

## Inferential results for LGBTQ 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

According to the inferential results, the p value from the chi-squared test is 0.000002266 which is less than 0.05, so we reject the null hypothesis. The odds ratio from the fisher's exact test is 4.27, so the vaccine has a 4.27 chance of being effective in stopping COVID. 

## Druggies

The null hypothesis: the vaccine doesn't exhibit its effectiveness for drug users. 

The alternative hypothesis: the vaccine does exhibit its effectiveness for drug users.

```{r}
DrugUser = FinalData[FinalData$DrugUser=='yes',]
head(DrugUser)
```


## Graphical results for Druggies

```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")
```

The Data shows that the vaccine is more likely to be less effective in drug users. The amount of people who received the vaccine and contracted COVID passed 80%, which is significantly higher than the amount of people who received the placebo and contracted COVID. 

## Numerical results for Druggies 
```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

We can see that 91.67% of the drug users who got COVID received the vaccine. This means that the vaccine is not effective in people who are drug users, because the amount of drug users who got the placebo and contracted COVID was 8.33%. 

## Inferential results for Druggies 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

We can see that the p value for the chi-squared test is 0.0000000001414 which is less than 0.05. That means we reject the null hypothesis. The odds ratio from the fisher's exact test is 0.087, which means that the vaccine has a lower chance of being effective for drug users. 

# Overall Results and Conclusions

First of all, we can conclude that the vaccine does exhibit effectiveness in preventing COVID in men. The numeric data shows that 26.20% men contracted COVID and got the vaccine compared to men who gotthe placebo which was 73.80%. The p-value from the chi-squared was 0.00000000000000022 which is less than 0.05 so we reject the null hypothesis, and the odds of the vaccine being effective in men is 2.8 so we can say that the vaccine is effective in preventing COVID in men. 

Second, We can say that the vaccine does exhibit effectiveness in stopping COVID in women. The numeric data shows us that 35.07% women who contracted COVID received the vaccine compared to the women who received the placebo which was 64.93%. Also, the p value  from the chi-squared test was 0.00000001914 which is less than 0.05 therefore we reject the null hypothesis. In addition, The odds of the vaccine being effective in women is about a 1.86 chance. Therefore, we can say the vaccine is effective in stopping COVID in women. 

Third, We can also say the vaccine is effective in stopping COVID in people who are LGBTQ. The numeric data shows the fact that 20% of people who are LGBTQ contracted COVID and also received the vaccine. The number of LGBTQ people who received the placebo was 80% which means the vaccine is effective in preventing COVID in people who are LGBTQ. The p-value from the chi-squared test is 0.000002266 which is less than 0.05, this means that we can reject the null hypothesis. On top of that, The odds ratio from the fisher's exact test is 4.27 which means the vaccine has a 4.27 chance of being effective in preventing COVID in people who are LGBTQ. Therefore, it is alright to say that that the vaccine is effective in stopping COVID in people who are LGBTQ. 

Finally, we can state that the vaccine is less likely to be effective in preventing COVID in drug users, because in the numeric data, approximately 91.67% of drug users who contracted COVID also received the vaccine compared to drug users who received the placebo which was 8.33%. The p-value from the chi-squared test was 0.0000000001414 which is less than 0.05, this means that we reject the null hypothesis, and the odds ratio from the fisher's exact test is 0.087 which means that the vaccine is less likely to effective in people who are drug users. Which means that we say that the vaccine is not effective in stopping COVID in drug users. 

Overall we can say that it is effective in stopping COVID in men, women, people who are LGBTQ, but not drug users. 

